---
title: "Charity Organization Dashboard"
author: "James Junior Baah"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: readable
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
```

## ðŸ“¥ Load Data

```{r load-data}
donors <- read_csv("Donors.csv")
beneficiaries <- read_csv("Beneficiaries.csv")
donations <- read_csv("Donations.csv")
events <- read_csv("Events.csv")
volunteers <- read_csv("Volunteers.csv")
event_participation <- read_csv("Event_Participation.csv")
```

## ðŸ” Data Exploration

```{r explore-data, message=FALSE}
glimpse(donors)
glimpse(donations)
glimpse(events)
```

## ðŸ’° Total Donations by Type

```{r donations-by-type}
donations %>%
  group_by(donation_type) %>%
  summarise(total = sum(amount)) %>%
  ggplot(aes(x = donation_type, y = total, fill = donation_type)) +
  geom_col() +
  labs(title = "Total Donations by Type", x = "Donation Type", y = "Amount") +
  theme_minimal()
```

## ðŸŽ’ Beneficiaries by Need

```{r beneficiaries-by-need}
beneficiaries %>%
  count(needs) %>%
  ggplot(aes(x = reorder(needs, n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Beneficiaries by Need", x = "Need", y = "Count") +
  theme_minimal()
```

## ðŸ™‹ Volunteers per Event

```{r volunteers-per-event}
event_participation %>%
  left_join(events, by = "event_id") %>%
  count(event_name) %>%
  ggplot(aes(x = reorder(event_name, n), y = n)) +
  geom_col(fill = "forestgreen") +
  coord_flip() +
  labs(title = "Volunteers per Event", x = "Event", y = "Volunteer Count") +
  theme_minimal()
```

## ðŸ“¤ Save a Plot

```{r save-plot, eval=FALSE}
# Example: Save the Donations by Type chart
# ggsave("donations_by_type.png", width = 6, height = 4)
```

## ðŸ’¡ Summary

This dashboard was built using R and `ggplot2` to visualize a SQL-backed charity database. It provides insights into donations, volunteers, and beneficiaries, and can be used to support decision-making and promote transparency.

---

_This is part of James Junior Baah's Data Analytics portfolio project. Connect with me on [LinkedIn](www.linkedin.com/in/james-kojo-junior-baah-881361336)._
